---
title: "NUC Dual Bayesian Inference Models for Bridgers et al. (2020)"
author: "Bianca Pfahl"
date: "19.02.2022"
output: 
  html_document:
    toc: true
    toc_depth: 4

---

```{r setup, include=FALSE}

## Packages from ABDA course

# package for convenience functions (e.g. ggplot2, dplyr, etc.)
library(tidyverse)

# package for Bayesian regression modeling
library(brms)

# package for visualization
library(tidybayes)
library(ggpubr)

# package to visualize 
library(bayesplot)

library(rstan)

library(xtable)

#devtools::install_github("michael-franke/aida-package")
library(aida)

# use the aida-theme for plotting
theme_set(theme_aida())

# global color scheme / non-optimized
project_colors = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#000000")

# setting theme colors globally
scale_colour_discrete <- function(...) {
  scale_colour_manual(..., values = project_colors)
}
scale_fill_discrete <- function(...) {
   scale_fill_manual(..., values = project_colors)
} 

```

#### 1. Data

All read in Data comes from Bridgers et. al (2020): Young children consider the expected utility of others’ learning to decide what to teach

Reading in data from the main experiment (experiment 1-3) and the norming study

```{r, include=FALSE}

# Reading in data from experiment 1 and relevant data labels

condition_labels <- 
  c("Rewards &\nCosts",
    "Different\nCosts",
    "Different\nRewards",
    "Medium Cost\nConflict",
    "High Cost\nConflict",
    "Extra-High Cost\nConflict"
  )

condition_labels_no_break <- 
  c("Rewards & Costs",
    "Different Costs",
    "Different Rewards",
    "Medium Cost Conflict",
    "High Cost Conflict",
    "Extra-High Cost Conflict"
  )

condition_levels <-
  c(
    "1: R&C", 
    "3: Diff. C", 
    "2: Diff. R", 
    "5: Med. Cost", 
    "4: RvC", 
    "7: 12 High High Cost"
  )

# Tidy data
choice_data <-
  read_csv('Data/EmpiricalData.csv', col_types = cols()) %>% 
  mutate(
    condition = 
      factor(Condition, levels = condition_levels, labels = condition_labels_no_break)
  )  %>%
  arrange(condition) %>% 
  mutate(condition_nr = c(1,2,3,4,5,6)) %>% 
  select(condition_nr, condition,red_toy = RedToy, sample_size = SampleSize)

choice_data

```


```{r, include=FALSE}

# Reading in Data from Experiment 2 & 3

# Experiment 2 data
d_exp2 <- read_csv('Data/exp2_data',  col_types = cols()) %>% 
  filter(condition=='Teach') %>% 
  select(subj_id, teach) %>% 
  mutate(condition = 'Exp2: Teach') %>% 
  select(condition, teach) %>% 
  add_count(teach) %>% 
  filter(teach=='red') %>% 
  distinct() %>% 
  mutate(sample_size = 25,
         condition_nr = 7) %>% 
  rename(red_toy = n) %>% 
  select(condition_nr, condition, red_toy, sample_size)
  
  
# Experiment 3 data
d_exp3 <- read_csv('Data/exp3_data',  col_types = cols()) %>% 
  filter(condition=='Exploration') %>% 
  select(subj_id, teach) %>% 
  mutate(condition = 'Exp3: Exploration') %>% 
  select(condition, teach) %>% 
  add_count(teach) %>% 
  filter(teach=='red') %>% 
  distinct() %>% 
  mutate(sample_size = 25,
         condition_nr = 8) %>% 
  rename(red_toy = n) %>% 
  select(condition_nr, condition, red_toy, sample_size)

all_choice_data <- bind_rows(choice_data, d_exp2, d_exp3)


```

Choice Data from Main Experiment:

```{r}
all_choice_data 
```

```{r, read in data, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}

# Reading in Norming Data

d_norm <- read_csv('Data/norm_data',  col_types = cols())

d_norm_tidy <- 
  d_norm %>%
  mutate(
    choice = factor(choice, levels = c("high_reward", "low_reward"))
    #gender = factor(gender, levels = c("male", "female"))
  ) %>% 
  mutate(Preference = if_else(choice=="high_reward", "lights", "music")) %>% 
  select(subj_id,choice,Preference)

# Prepare Data from the Norming Study

d_norm_compact <- 
  d_norm_tidy %>% 
  count(Preference) %>% 
  filter(Preference=="lights") %>% 
  mutate(SampleSize = nrow(d_norm_tidy),
         Lights = n) %>% 
  select(Preference,Lights, SampleSize)


```

Norming Data:

```{r}
d_norm_compact
```

Combining all data from main experiment & norming study into a list to parse to Stan

cost matrix: dimension 1 stands for: activation [1] or discovery [2] cost, dimension 2 stands for the 4 toy difficulty levels [1-4]

reward_vec_red/yellow and cost_vec_red/yellow indicate the indices to access the correct rewards & costs per condition

```{r}

# transforming cost values from (0,100) interval to (0,1) inverval
costs = matrix(c(1,1,1,4,2,29,2,79), nrow=2, ncol=4)
costs = costs / 100


choice_data_list = list(
  k = all_choice_data$red_toy, n = all_choice_data$sample_size, n_conditions = 8, rewards_vec_red = c(1,1,2,2,2,2,2,1),
  rewards_vec_yellow = c(2,1,1,1,1,1,1,1), cost_vec_red = c(1,1,1,1,1,1,1,1), cost_vec_yellow  = c(3,3,1,2,3,4,3,3),costs = costs
)

d_norm_list <- list(k_norm = d_norm_compact$Lights, n_norm = d_norm_compact$SampleSize)

# Combining data from norming study and rest

data_and_norm <- append(choice_data_list, d_norm_list)

data_and_norm

# creating a slightly altered version for the cost models

```

```{r, warning=FALSE, message=FALSE, echo=FALSE, include=FALSE}

# Combining all data from main experiment & norming study into a list to parse to Stan for the Cost Model

choice_data_list_cost_model = list(
  k = all_choice_data$red_toy, n = all_choice_data$sample_size, n_conditions = 8, rewards_vec_red = c(1,1,2,2,2,2,2,1),
  rewards_vec_yellow = c(2,1,1,1,1,1,1,1), cost_vec_red = c(1,1,1,1,1,1,1,1), cost_vec_yellow  = c(3,3,1,2,3,4,3,3)
)

d_norm_list_cost_model <- list(k_norm = d_norm_compact$Lights, n_norm = d_norm_compact$SampleSize)

# Combining data from norming study and rest

data_and_norm_cost_model <- append(choice_data_list_cost_model, d_norm_list_cost_model)
```

The following code chunk generates the data for Table 1 from the thesis:

```{r}

pretty_costs <- tibble(
  Costs = c("low", "medium", "high", "extra high"),
  Activation_Cost = c(0.01,0.01,0.02,0.02),
  Discovery_Cost = c(0.01,0.04,0.29,0.79)
)

pretty_costs
print(xtable(pretty_costs, type = "latex"), file = "costs_table.tex")

```

#### 2. Models

#### 2.1 Main Model: Dual Bayesian Inference Model

```{stan, output.var="Dual_Bayesian_Inference_Model", eval=F}


// Dual Bayesian Inference Model (Main Model)
// The model performs dual Bayesian inference over two distinct data sets: the main experiment data and norming study
// data from Bridgers et al. (2020)

data { 
  int<lower=1> n_conditions;
  int<lower=0> k[n_conditions];
  int<lower=1> n[n_conditions];
  int<lower=1> n_norm;
  int<lower=0> k_norm;
  
  // REWARD indices
  int rewards_vec_red[n_conditions]; 
  int rewards_vec_yellow[n_conditions]; 
  
  // COST indices
  int cost_vec_red[n_conditions];
  int cost_vec_yellow[n_conditions];
  
  // COST VECTOR
  real costs[2,4];
} 
parameters {
  
  //REWARD VECTOR: reward[1] corresponds to the music mechanism, reward[2] corresponds to the light mechanism
  real<lower=0, upper=1> rewards[2];
  
} 

model {
  real utility_toy_red;
  real utility_toy_yellow;
  real prob_choice_red;
  real prob_lights;
  
  // HYPERPARAMETERS
  real alpha = 5;
  real explo_para = 0.5;
  
  // PRIORS 
  rewards[1]  ~ beta(1,1);
  rewards[2] ~ beta(1,1);
  
 // MAIN LOOP: going through all 6 conditions of the experiment
  
  for (i in 1:n_conditions) {
    
    // UTILITIES
    // computing utility values for red toy and yellow toy respectively, according to condition
    utility_toy_red = rewards[rewards_vec_red[i]] - costs[1][cost_vec_red[i]] + explo_para * rewards[rewards_vec_yellow[i]] - costs[2][cost_vec_yellow[i]];
    utility_toy_yellow = rewards[rewards_vec_yellow[i]] - costs[1][cost_vec_yellow[i]] + explo_para * rewards[rewards_vec_red[i]] - costs[2][cost_vec_red[i]];
  
    // PROBABILITY FOR RED TOY
    // computing the probability for a child to choose the red toy using softmax choice rule
    prob_choice_red = exp(alpha * utility_toy_red) / (exp(alpha * utility_toy_red) + exp(alpha * utility_toy_yellow));
    
    // POSTERIOR
    // Observed amount of decisions for red toy
    k[i] ~ binomial(n[i], prob_choice_red);
  }
  
  // NORMING STUDY
  
  //PROBABILITY FOR LIGHT TOY
  // computing the probability for a child to choose the light toy using softmax choice rule
  prob_lights = exp(alpha * rewards[2]) / (exp(alpha * rewards[2]) + exp(alpha*rewards[1])); 
  
  // NORMING STUDY POSTERIOR 
  // Observed amount of decisions for light toy
  k_norm ~ binomial(n_norm, prob_lights);
}

```

```{r, echo=FALSE, results=FALSE}

# running the Stan model

Dual_Bayesian_Inference_Model <- stan(file='Models/Dual_Bayesian_Inference_Model.stan',   
                data=data_and_norm,
                iter=10000,
                chains=4,
                thin=1
)

```

The following code chunk creates the data for Table 2 from the thesis:

```{r}

Dual_Bayesian_Inference_Model_tibble <- summary(Dual_Bayesian_Inference_Model)$summary

Dual_Bayesian_Inference_Model_tibble
print(xtable(Dual_Bayesian_Inference_Model_tibble, type = "latex"), file = "Dual_Bayesian_Inference_Model_Output.tex")

```

Visualising the Results: Dual Bayesian Inference Model

The following code chunk creates Figure 6 (Left) from the thesis:

```{r, WARNING=FALSE}

# extract samples from the stanfit object 
Dual_Bayesian_Inference_Model_samples <- tidybayes::tidy_draws(Dual_Bayesian_Inference_Model)

Dual_Bayesian_Inference_Model_samples_matrix <- as.matrix(Dual_Bayesian_Inference_Model_samples)

mcmc_areas(Dual_Bayesian_Inference_Model_samples_matrix,
           regex_pars = "rewards\\[[1-2]\\]", prob = 0.8, area_method = "equal height") +
  scale_y_discrete(
      labels = c("rewards[1]" = "Reward - Music",
             "rewards[2]" = "Reward - Lights"
             )) +
       labs(x="Density",
   title = "Posterior Distributions: Rewards",
   subtitle = "with medians and 80% intervals") +
  theme(plot.title.position = "plot",
        plot.caption.position =  "plot") +
  legend_move("none")

dev.copy(png,filename="Dual_Bayesian_Inference_Model_posteriors.png")
dev.off ()

```

The following code chunk creates Figure 6 (Right) from the thesis:

```{r}

Dual_Bayesian_Inference_Model_samples <- Dual_Bayesian_Inference_Model_samples %>% 
                  mutate(difference = `rewards[2]` - `rewards[1]`,
                         mean_difference = mean(difference))

# accessing the mean difference between the two rewards
mean_difference = Dual_Bayesian_Inference_Model_samples$mean_difference[1]

color_scheme_set("teal")

Dual_Bayesian_Inference_reward_difference_plot <- ggplot(Dual_Bayesian_Inference_Model_samples, aes(x=`difference`)) + geom_density(color="#69b3a2",fill="mediumaquamarine", alpha=0.5) + geom_vline(xintercept=mean_difference, color="violetred4", size=1,linetype="dashed") +
  labs(x=expression(bold(R[light] - R[music])), y="Density",
   title = "Difference in Reward",
   subtitle = expression(R[light] - R[music])) +
  theme(plot.title.position = "plot",
        plot.caption.position =  "plot")

Dual_Bayesian_Inference_reward_difference_plot

dev.copy(png,filename="Dual_Bayesian_Inference_Reward_Difference.png")
dev.off ()

```


Hypothesis Tests: Dual Bayesian Inference Model

The following code chunk creates the data for Table 3 from the thesis:

```{r}

# employing a one-sided hypothesis from the brms package to check if the posterior for reward music and reward light are credibly different

hypothesis_test_Dual_Bayesian_Inference_Model <- hypothesis(Dual_Bayesian_Inference_Model_samples, 'rewards[2] > rewards[1]')

hypothesis_test_Dual_Bayesian_Inference_Model

print(xtable(hypothesis_test_Dual_Bayesian_Inference_Model$hypothesis, type = "latex"), file = "hypothesis_table_2rewards.tex")


```

##### 2.2 Four Rewards Model

Stan Model:

```{stan, output.var="Four_Rewards_Model", eval=F}

// Four Rewards Model
// The model performs dual Bayesian inference over two distinct data sets: the main experiment data and norming study
// data from Bridgers et al. (2020) and estimates separat values per activation and discovery reward per toy. 

data { 
  int<lower=1> n_conditions;
  int<lower=0> k[n_conditions];
  int<lower=1> n[n_conditions];
  int<lower=1> n_norm;
  int<lower=0> k_norm;
  
  // REWARDS indices
  int rewards_vec_red[n_conditions]; 
  int rewards_vec_yellow[n_conditions]; 
  
  // COSTS indices
  int cost_vec_red[n_conditions];
  int cost_vec_yellow[n_conditions];
  
  // COST VECTOR
  real costs[2,4];
} 
parameters {
  
 //REWARD VECTOR: reward[1] corresponds to the music mechanism, reward[2] corresponds to the light mechanism
  real<lower=0, upper=1> rewards[2,2];

} 

model {
  real utility_toy_red;
  real utility_toy_yellow;
  real prob_choice_red;
  real prob_lights;
  
  // HYPERPARAMETERS
  real alpha = 5;
  real explo_para = 0.5;
  
  // REWARDS
  
  // activation reward for music effect toy
  rewards[1][1]  ~ beta(1,1);
   
  // activation reward for light effect toy
  rewards[1][2] ~ beta(1,1);
  
  // discovery reward for music effect toy
  rewards[2][1] ~ beta(1,1);
  
  // discovery reward for light effect toy
  rewards[2][2] ~ beta(1,1);
  
 // MAIN LOOP: going through all 6 conditions of the experiment
  
  for (i in 1:n_conditions) {
    
    // UTILITIES
    // computing utility values for red toy and yellow toy respectively according to condition, using seperate
    // variables for discovery and action reward depending on context the toy is presented in
    utility_toy_red = rewards[1][rewards_vec_red[i]] - costs[1][cost_vec_red[i]] + explo_para * rewards[2][rewards_vec_yellow[i]] -   costs[2][cost_vec_yellow[i]];
    utility_toy_yellow = rewards[1][rewards_vec_yellow[i]] - costs[1][cost_vec_yellow[i]] + explo_para * rewards[2][rewards_vec_red[i]] - costs[2][cost_vec_red[i]];
  
    // PROBABILITIES
   // computing the probability for a child to choose the red toy using softmax choice rule
    prob_choice_red = exp(alpha * utility_toy_red) / (exp(alpha * utility_toy_red) + exp(alpha * utility_toy_yellow));
    
    // POSTERIOR
    // Observed amount of decisions for red toy
    k[i] ~ binomial(n[i], prob_choice_red);
  }
  
  // NORMING STUDY
  
  //PROBABILITY FOR LIGHT TOY
  // computing the probability for a child to choose the light toy using softmax choice rule
  prob_lights = exp(alpha * rewards[2][2]) / (exp(alpha * rewards[2][2]) + exp(alpha*rewards[2][1]));
  
  // NORMING STUDY POSTERIOR 
  // Observed amount of decisions for light toy
  k_norm ~ binomial(n_norm, prob_lights);
}


```


```{r, echo=FALSE, results=FALSE}

Four_Rewards_Model <- stan(file='Models/Four_Rewards_Model.stan',   
                data=data_and_norm,
                iter=10000,
                chains=4,
                thin=1
)

```

The following code chunk creates the data for Table 4 from the thesis:

```{r}
Four_Rewards_Model
Four_Rewards_Model_tibble <- summary(Four_Rewards_Model)$summary
print(xtable(Four_Rewards_Model_tibble, type = "latex"), file = "four_rewards_summary.tex")


```


The following code chunk creates the data for Table 5 from the thesis:

```{r}

# is discovery reward > activation reward for music toy?
Four_Rewards_Model_hypothesis1 <- hypothesis(Four_Rewards_Model, "rewards[2,1] > rewards[1,1]") 

# is discovery reward > activation reward for light toy?
Four_Rewards_Model_hypothesis2 <- hypothesis(Four_Rewards_Model, "rewards[2,2] > rewards[1,2]")

Four_Rewards_Model_Hypothesis_Tests <- bind_rows(Four_Rewards_Model_hypothesis1$hypothesis,Four_Rewards_Model_hypothesis2$hypothesis)

Four_Rewards_Model_Hypothesis_Tests

print(xtable(Four_Rewards_Model_Hypothesis_Tests, type = "latex"), file = "Four_Rewards_Model_Hypothesis_Tests.tex")

```

The following code chunk creates Figure 8 (Left) from the thesis:

```{r, WARNING=FALSE, results=FALSE}
# extract samples from the stanfit object
Four_Rewards_Model_samples <- tidybayes::tidy_draws(Four_Rewards_Model)

Four_Rewards_Model_samples_matrix <- as.matrix(Four_Rewards_Model_samples)

Four_Rewards_Model_Posteriors_plot <- mcmc_areas(Four_Rewards_Model_samples_matrix,
           regex_pars = "rewards\\[[1-2],[1-2]\\]", prob = 0.5, prob_outer = 0.95,  area_method = "equal heigh") +
  scale_y_discrete(
      labels = c("rewards[1,1]" = "Activation Reward - Music",
             "rewards[1,2]" = "Activation Reward - Lights",
             "rewards[2,1]" = "Discovery Reward - Music",
             "rewards[2,2]" = "Discovery Reward - Lights")) +
      labs(x="Density",
      title = "Posterior Distributions: Rewards",
      subtitle = "with medians and 80% intervals") +
      theme(plot.title.position = "plot",
        plot.caption.position =  "plot")

Four_Rewards_Model_Posteriors_plot 

dev.copy(png,filename="Four_Rewards_Model_Posteriors.png")
dev.off ()

```

The following code chunk creates Figure 8 (Right) from the thesis:

```{r, results=FALSE}

Four_Rewards_Model_samples_matrix <- Four_Rewards_Model_samples %>% 
    mutate(R_Act_Music = Four_Rewards_Model_samples$`rewards[1,1]`,
           R_Act_Light = Four_Rewards_Model_samples$`rewards[1,2]`,
           R_Dis_Music = Four_Rewards_Model_samples$`rewards[2,1]`,
           R_Dis_Light = Four_Rewards_Model_samples$`rewards[2,2]`)

Four_Rewards_Model_Scatter_Plots <- mcmc_pairs(Four_Rewards_Model_samples_matrix, pars = c("R_Dis_Light", "R_Act_Light", "R_Dis_Music" ,"R_Act_Music"),
                             off_diag_args = list(size = 1, alpha = 0.25))

Four_Rewards_Model_Scatter_Plots

dev.copy(png,filename="Four_Rewards_Model_Plots.png")
dev.off ()

```

      
#### 2.3 Costs Model

Cost Model in Stan:

```{stan, output.var="Cost_Model", eval=F}

// Cost Model
// The model performs dual Bayesian inference over two distinct data sets: the main experiment data and norming study
// data from Bridgers et al. (2020) and estimates separate cost values per discovery and activation costs as well as
// per toy difficulty level.

data { 
  int<lower=1> n_conditions;
  int<lower=0> k[n_conditions];
  int<lower=1> n[n_conditions];
  int<lower=1> n_norm;
  int<lower=0> k_norm;
  
  // REWARDS indices
  int rewards_vec_red[n_conditions]; 
  int rewards_vec_yellow[n_conditions]; 
  
  // COSTS indices
  int cost_vec_red[n_conditions];
  int cost_vec_yellow[n_conditions];
  
  
} 
parameters {
  
  // COSTS
  real<lower=0, upper=1> costs[2,4];
  
} 

model {
  real utility_toy_red;
  real utility_toy_yellow;
  real prob_choice_red;
  real prob_lights;
  
  // HYPERPARAMETERS
  real alpha = 5;
  real explo_para = 0.5;
  
  // REWARDS
  real rewards[2]  = {0.35, 0.65};
  
  // PRIORS 
  
  for (i in 1:2) {
    for (j in 1:4){
      costs[i][j] ~ beta(1,1);
    }
  }
  
 // MAIN LOOP: going through all 6 conditions of the experiment
  
  for (i in 1:n_conditions) {
    
    // UTILITIES
    // computing utility values for red toy and yellow toy respectively, according to condition, using eight separately estimated
    // cost values
    utility_toy_red = rewards[rewards_vec_red[i]] - costs[1][cost_vec_red[i]] + explo_para * rewards[rewards_vec_yellow[i]] - costs[2][cost_vec_yellow[i]];
    utility_toy_yellow = rewards[rewards_vec_yellow[i]] - costs[1][cost_vec_yellow[i]] + explo_para * rewards[rewards_vec_red[i]] - costs[2][cost_vec_red[i]];
  
    // PROBABILITY FOR RED TOY
    // computing the probability for a child to choose the red toy using softmax choice rule
    prob_choice_red = exp(alpha * utility_toy_red) / (exp(alpha * utility_toy_red) + exp(alpha * utility_toy_yellow));
    
    // POSTERIOR
    // Observed amount of decisions for red toy
    k[i] ~ binomial(n[i], prob_choice_red);
  }
  
   // NORMING STUDY
  
  //PROBABILITY FOR LIGHT TOY
  // computing the probability for a child to choose the light toy using softmax choice rule
  prob_lights = exp(alpha * rewards[2]) / (exp(alpha * rewards[2]) + exp(alpha*rewards[1]));
  
  // NORMING STUDY POSTERIOR 
  // Observed amount of decisions for light toy
  k_norm ~ binomial(n_norm, prob_lights);
}


```


```{r, echo=FALSE, results=FALSE}

# running the Stan model

Cost_Model <- stan(file='Models/Cost_Model.stan',   
                data=data_and_norm_cost_model,
                iter=5000,
                chains=4,
                thin=1
)

```

The following code chunk displays the data for Table 6 from the thesis:

```{r}

cost_stanfit_tibble <- summary(Cost_Model)$summary
print(xtable(cost_stanfit_tibble, type = "latex"), file = "Cost_Model_Output.tex")

Cost_Model

```

Hypothesis Tests for the Cost Model

Activation Costs

The following code chunk displays the data for Table 7 from the thesis:

```{r}

Cost_Model_samples <- tidybayes::tidy_draws(Cost_Model)

# hypothesis tests for activation costs

cost_l1l2 <- hypothesis(Cost_Model_samples, "costs[1,1] > costs[1,2]") 

cost_l2l3 <- hypothesis(Cost_Model_samples, "costs[1,2] > costs[1,3]")

costl3l4 <- hypothesis(Cost_Model_samples, "costs[1,3] > costs[1,4]")

cost_extra <- hypothesis(Cost_Model_samples, "costs[1,1] > costs[1,4]")

Cost_Model_Hypothesis_Tests <- bind_rows(cost_l1l2$hypothesis,cost_l2l3$hypothesis,costl3l4$hypothesis,cost_extra$hypothesis)

print(xtable(Cost_Model_Hypothesis_Tests, type = "latex"), file = "Cost_Model_Activation_Cost_Hypothesis_Tests.tex")

Cost_Model_Hypothesis_Tests

```

Additional Hypothesis Tests on the Cost Model's discovery costs (not illustrated in a table in the thesis)

```{r}

dis_cost_l1l2 <- hypothesis(Cost_Model_samples, "costs[2,1] < costs[2,2]") 

dis_cost_l2l3 <- hypothesis(Cost_Model_samples, "costs[2,2] < costs[2,3]")

dis_costl3l4 <- hypothesis(Cost_Model_samples, "costs[2,3] < costs[2,4]")

table_all_hypothesis_tests_dis_cost <- bind_rows(dis_cost_l1l2$hypothesis,dis_cost_l2l3$hypothesis,dis_costl3l4$hypothesis)

table_all_hypothesis_tests_dis_cost

print(xtable(table_all_hypothesis_tests_dis_cost, type = "latex"), file = "Cost_Model_Discovery_Costs_Hypothesis_Tests.tex")

```

The following code chunk creates Figure 10 from the thesis:

```{r, results=FALSE}

Cost_Model_samples_matrix <- as.matrix(Cost_Model_samples)

costs_sample_plot_activation <- mcmc_areas(Cost_Model_samples_matrix,
      regex_pars = "costs\\[[1],[1-4]\\]", prob = 0.8) +
      scale_y_discrete(
      labels = c("costs[1,1]" = "Low",
             "costs[1,2]" = "Medium",
             "costs[1,3]" = "High",
             "costs[1,4]" = "Extra High")) +
       labs(x="Density",
   title = "Posterior Distribution: Activation Costs",
   subtitle = "with medians and 80% intervals") +
  theme(plot.title.position = "plot",
        plot.caption.position =  "plot")

color_scheme_set("purple")

costs_sample_plot_discovery <- mcmc_areas(Cost_Model_samples_matrix,
      regex_pars = "costs\\[[2],[1-4]\\]", prob = 0.8) +
      scale_y_discrete(
      labels = c(
             "costs[2,1]" = "Low",
             "costs[2,2]" = "Medium",
             "costs[2,3]" = "High",
             "costs[2,4]" = "Extra High")) +
       labs(x="Density",
   title = "Posterior Distributions: Discovery Costs",
   subtitle = "with medians and 80% intervals") +
  theme(plot.title.position = "plot",
        plot.caption.position =  "plot")

costs_sample_plot_activation

dev.copy(png,filename="activation_costs.png")
dev.off ()

costs_sample_plot_discovery

dev.copy(png,filename="discovery_costs.png")
dev.off ()

```

#### 2.4 Discovery Difficulty Model

Stan Model:

```{stan, output.var="Discovery_Difficulty_Model", eval=F}

// Discovery Difficulty Model
// The model performs dual Bayesian inference over two distinct data sets: the main experiment data and norming study
// data from Bridgers et al. (2020) and estimates four separate levels for exploration certainty parameter

data { 
  int<lower=1> n_conditions;
  int<lower=0> k[n_conditions];
  int<lower=1> n[n_conditions];
  int<lower=1> n_norm;
  int<lower=0> k_norm;
  
  // REWARDS indices
  int rewards_vec_red[n_conditions]; 
  int rewards_vec_yellow[n_conditions]; 
  
  // COSTS indices
  int cost_vec_red[n_conditions];
  int cost_vec_yellow[n_conditions];
  
  // COSTS
  real<lower=0, upper=1> costs[2,4];
  
} 
parameters {
  
  // Exploration Parameter PE (one per toy difficulty level)
  real<lower=0, upper=1> explo_para[4];
} 

model {
  real utility_toy_red;
  real utility_toy_yellow;
  real prob_choice_red;
  real prob_lights;
  
  // HYPERPARAMETER
  real alpha = 5;
  
  // REWARDS
  real rewards[2]  = {0.35, 0.65};
  
  // PRIOR
  for (i in 1:4) {
      explo_para[i] ~ beta(1,1);
  }
  
 // MAIN LOOP: going through all 6 conditions of the experiment
  
  for (i in 1:n_conditions) {
    
    // UTILITIES
    // computing utility values for red toy and yellow toy respectively, according to condition, with four separate exploration
    // certainty parameter settings depending on toy difficulty level 
    utility_toy_red = rewards[rewards_vec_red[i]] - costs[1][cost_vec_red[i]] + explo_para[cost_vec_yellow[i]] * rewards[rewards_vec_yellow[i]] - costs[2][cost_vec_yellow[i]];
    utility_toy_yellow = rewards[rewards_vec_yellow[i]] - costs[1][cost_vec_yellow[i]] + explo_para[cost_vec_red[i]] * rewards[rewards_vec_red[i]] - costs[2][cost_vec_red[i]];
  
     // PROBABILITY FOR RED TOY
    // computing the probability for a child to choose the red toy using softmax choice rule
    prob_choice_red = exp(alpha * utility_toy_red) / (exp(alpha * utility_toy_red) + exp(alpha * utility_toy_yellow));
    
    // POSTERIOR
    // Observed amount of decisions for red toy
    k[i] ~ binomial(n[i], prob_choice_red);
  }
  
  // NORMING STUDY
  
  //PROBABILITY FOR LIGHT TOY
  // computing the probability for a child to choose the light toy using softmax choice rule
  prob_lights = exp(alpha * rewards[2]) / (exp(alpha * rewards[2]) + exp(alpha*rewards[1]));
  
  // NORMING STUDY POSTERIOR 
  // Observed amount of decisions for light toy
  k_norm ~ binomial(n_norm, prob_lights);
}


```


```{r, results=FALSE}

# running the Stan model

Discovery_Difficulty_Model <- stan(file='Models/Discovery_Difficulty_Model.stan',   
                data=data_and_norm,
                iter=10000,
                chains=4,
                thin=1,
)

```

The following code chunk displays the data for Table 8:

```{r}

Discovery_Difficulty_Model

Discovery_Difficulty_Model_tibble <- summary(Discovery_Difficulty_Model)$summary
print(xtable(Discovery_Difficulty_Model_tibble, type = "latex"), file = "discovery_difficulty_model_output.tex")

```

Hypothesis Tests for the Discovery Difficulty Model

The following code chunk displays the data for Table 9 from the thesis:

```{r}

Discovery_Difficulty_Model_hypothesis1 <- hypothesis(Discovery_Difficulty_Model, "explo_para[1] > explo_para[2]")
Discovery_Difficulty_Model_hypothesis2 <- hypothesis(Discovery_Difficulty_Model, "explo_para[3] > explo_para[2]")
Discovery_Difficulty_Model_hypothesis3 <- hypothesis(Discovery_Difficulty_Model, "explo_para[4] > explo_para[3]")

Discovery_Difficulty_Model_all_hypotheses <- bind_rows(Discovery_Difficulty_Model_hypothesis1$hypothesis,         Discovery_Difficulty_Model_hypothesis2$hypothesis,Discovery_Difficulty_Model_hypothesis3$hypothesis)

Discovery_Difficulty_Model_all_hypotheses

print(xtable(Discovery_Difficulty_Model_all_hypotheses, type = "latex"), file = "Discovery_Difficulty_Model_Hypothesis_Tests.tex")



```

The following code chunk creates Figure 12 from the thesis:

```{r, results=FALSE}
Discovery_Difficulty_Model_samples <- tidybayes::tidy_draws(Discovery_Difficulty_Model)


Discovery_Difficulty_Model_samples_plus_differences <- Discovery_Difficulty_Model_samples %>% 
                  mutate(`difference_level1_level2` = `explo_para[1]` - `explo_para[2]`,
                         mean_l1l2 = mean(difference_level1_level2),
                         difference_level2_level3 = `explo_para[2]` - `explo_para[3]`,
                         mean_l2l3 = mean(difference_level2_level3),
                         difference_level3_level4 = `explo_para[3]` - `explo_para[4]`,
                         mean_l3l4 = mean(difference_level3_level4))


mean_l1l2 <- Discovery_Difficulty_Model_samples_plus_differences$mean_l1l2

Discovery_Difficulty_Model_Differences_l1l2 <- ggplot(Discovery_Difficulty_Model_samples_plus_differences, aes(x=`difference_level1_level2`)) + geom_density(color="goldenrod3",fill="orange", alpha=0.5) + geom_vline(xintercept=mean_l1l2, color="violetred4", size=1,linetype="dashed") + labs(x=bquote(bold(P[E] ~ "low" ~ - P[E] ~ "medium")), y="Density") +
  theme(axis.title = element_text(size = 11)) + xlim(-1, 0.6)

mean_l2l3 <- Discovery_Difficulty_Model_samples_plus_differences$mean_l2l3

Discovery_Difficulty_Model_Differences_l2l3 <- ggplot(Discovery_Difficulty_Model_samples_plus_differences, aes(x=`difference_level2_level3`)) + geom_density(color="tan2",fill="tan1", alpha=0.5) + geom_vline(xintercept=mean_l2l3, color="violetred4", size=1,linetype="dashed") + labs(x=bquote(bold(P[E] ~ "medium" ~ - P[E] ~ "high")), y="Density") + theme(axis.title = element_text(size = 11)) + xlim(-1, 0.6)

mean_l3l4 <- Discovery_Difficulty_Model_samples_plus_differences$mean_l3l4

Discovery_Difficulty_Model_Differences_l3l4 <- ggplot(Discovery_Difficulty_Model_samples_plus_differences, aes(x=`difference_level3_level4`)) + geom_density(color="tomato4",fill="tomato3", alpha=0.5) + geom_vline(xintercept=mean_l3l4, color="violetred4", size=1,linetype="dashed") + labs(x=bquote(bold(P[E] ~ "high" ~ - P[E] ~ "extra high")), y="Density") + theme(axis.title = element_text(size = 11)) + xlim(-1, 0.6)

Discovery_Difficulty_Model_Posteriors <- ggpubr::ggarrange(Discovery_Difficulty_Model_Differences_l1l2, Discovery_Difficulty_Model_Differences_l2l3, Discovery_Difficulty_Model_Differences_l3l4) + 
  labs(x=bquote(bold(P[E] ~ "high" ~ - P[E] ~ "extra high")), y="Density",
   title = "Difference in Exploration Certainty Parameter",
   subtitle = bquote("Posterior Densities (ordered)")) +
  theme(plot.title.position = "plot",
        plot.caption.position =  "plot")

Discovery_Difficulty_Model_Posteriors <- ggpubr::annotate_figure(
  Discovery_Difficulty_Model_Posteriors,
  top = text_grob("Difference in Exploration Certainty Parameter",
  size= 14,
  face="bold")
) 

Discovery_Difficulty_Model_Posteriors

dev.copy(png,filename="PE_differences.png")
dev.off ()
```


